### 2022-11-24 manuscript.Rmd
24 nov

---
title: 'Quantifying unmet ICU beds in tertiary hospitals in urban India using Propensity
  Score Matching '
author: "Maria Eriksson"
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
table_nums <- captioner::captioner(prefix = "Tab.")
```

<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->

<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->

<!-- Structure and most comments are from the Strengthening the
reporting of observational studies in epidemiology (STROBE) statement,
see https://www.strobe-statement.org/. For more explanations see the
paper Strengthening the Reporting of Observational Studies in
Epidemiology (STROBE):explanation and elaboration.  -->

Abbreviations
============
ACS - American college of surgeons

AUC - Area Under Curve

ER - Emergency room

GCS - Glasgow coma scale

HIC - High-income countries

ISS - Injury Severity score

LMIC - Low- and middle-income countries

NISS - New Injury Severity Score

PSM - propensity score matching

RCT - Randomized case-control study

TITCO -Towards Improved Trauma Care Outcomes

TQIP - Trauma Care Improvement Program

<!-- Indicate the study’s design with a commonly used term in the
title, for example: "Associations between resuscitation interventions
and opportunities for improvement in adult trauma patients: A cohort
study" -->

Abstract
========

<!-- Provide in the abstract an informative and balanced summary of
what was done and what was found. Not more than 300 words. -->

Background
----------
Trauma is a big cuse of mortality and morbidity in India, and it is a leading cause of death among young people. It has long been suggested that ICU resources are insufficient and that this is one of the factors that lead to Indian in-hospital motality being higher than in high-income countries.

Methods
-------
Propensity score matcing shall be applied to the TITCO dataset containing 16000 trauma patient details, in an effort to identify patients who did not go to the ICU but met the same treatment inclusion criteria as the patients that did receive ICU treatment, and as such reasonably should have received ICU treatment.

Results
-------

Conclusion
----------

Introduction
============

<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->

Trauma globally and in India
----------------------------
Trauma is a substantial global health problem that disproportionately affects low and middle-income countries (LMIC), resulting in significant mortality and morbidity @traumaLMIC. 4.8 million people die from trauma annually, 90% are in LMIC and 20% in India alone @onemilliondead. Despite having only 1% of the global vehicle fleet, a disproportionate 10% of global traffic casualties occur in India @indiantrauma. Half of trauma mortality in India occur in-hospital, but this rate varies greatly by region and mode of injury @indiantrauma @nationalcrime @titco . There is no national trauma registry in India, but isolated studies have shown that in-hospital mortality within 30 days of hospitalization is higher than similarly injured patients in HIC hospitals @indiantrauma. Globally, an estimated two million lives could be saved annually, if the appropriate and timely trauma care provided by hospitals in many high-income countries were available and provided in low to-middle-income countries (LMICs) @indiantrauma.

Health care professionals in many LMIC testify that their medical assessment oftentimes is that a trauma patient should be admitted to the ICU, but are not because such resources are scarce, overloaded and unavailable. Their medical judgement is that the lives of these patients were salvageable, and that these patients died as a direct result of not being admitted to the ICU @stuffLMICdoctorssay.

Trauma is a main cause of mortality and morbidity to otherwise healthy and young individuals, and dispropotionately affect this demographic. Most casualties affected the age group 30-45 years, followed by the age bracket 18-30 years @nationalcrime. People in this demographic are also oftentimes the primary or sole breadwinners for their families. Therefore the detrimental impact of trauma goes beyond mortality or substantial loss of disease-free years of life for the affected individuals, it also carries a considerable financial and social harm on their family members and communities @traumaLMIC @nationalcrime.

A diverse mix of public, for-profit and not-for-profit private actors provide health services in India, where private providers dominate. Public sector make up 20% of health care expenditure, and this 20% make up 1% of GDP. Both the proportion of public sector spending and the proportion of GDP is among the lowest in the world @goodstats. According to a WHO 2022 report, India has 76 709 non-ventilator supported and 39 476 ventilator-supported ICU beds @ICUbeds, out of which 10% of intitutions and 30% of ICU beds are public and financially feasibile for many Indians @indianicu @goodstats. Public hospitals have more or less perpetual shortage of ICU beds. If it is the attending physicians assessment that ICU care is medically necessary a patient or their relatives may choose to go to a private hospital, if the transfer is medically advisable. For a majority of patients it is financially unfeasible to pay out of pocket for the specialized ambulance or private hospitalization, that cost approx 100 times the average per capita income @perpetual @indianicu. Because of the type and severity of injuries, ICU resources are often essential in care of trauma patients @titco.

<!-- KommentarJohanna: I suggest adding more details about the context in India, the health care system, how common trauma is in India, how many ICU beds there are, why ICU care is important for the trauma population and so on.. 
-->
<!-- KommentarMartin:You can definitely write more about the health care system in India, the trauma system, the availability of ICUs, how common ICU care is in trauma, why trauma patients need ICU care, how long they stay, outcomes etc -->

<!-- Respons: Ja det är magert men det finns en orsak till det, jag har undvikit detta av två skäl
 1.Datan som står att finna om ICU vård resurserna i Indien är lite spretig, men enligt WHO så är den runt 8 per 100.000 och andra trovärdiga bedömare är där nånstans +- 50%. Det är i paritet med Sverige? Förklaringen står kanske att finna i att definitionen av en ICU bed inte är vedertagen men trots mycket läsande innantill har jag inte kunnat avfärda eller förklara detta. 
2. pkt 1 lyfter en fråga som ökar komplexiteten: det finns lite olika levels av ICU beds, vilken är det vi menar? Eftersom statistiken på ICU resurser är dåligt differentierad och TITCO datasetet inte specificerar det, ska frågan undvikas för att inte öka komplexiteten för sig själv? -->

Methods to examine trauma care and ensure equivalence
-----------------------------------------------------
In order to improve and ensure equivalence of intra-hospital care, trauma quality improvement programs (TQIP) have been developed. The pilot and most notable example is the TQIP developed by the American College of Surgeons ACS. Since the implementation of the ACS pilot, TQIP has evolved to a broad term that include many kinds of processes to improve the quality of trauma care, such as mortality and morbidity conferences, using audit filters or comparing risk adjusted outcomes over time and between centres. The aim is to find discrepancies in treatment for patients presenting with similar conditions. This difference in in-hospital mortality is likely multi-factorial, and any factors have been proposed to explain it such as differences in triage and delay-of-care, as well as shortage of hospital resources including ICU beds @onemilliondead ,@traumaLMIC . 
TQIPs aim to ensure patients get equivalent care within and across hospitals, to benchmark performance of health care providers versus their peers and to improve trauma care by developing best practice guidelines @tqip.

TQIP often originate in high-income countries (HIC), they are neither applicable nor easily transferred to the hospital conditions of LMIC @frailty @gunshot  @onemilliondead. For trauma patients, there exist significant structural differences between HIC and LMIC. These differences are in factors such as injury mechanisms and severity, health care resources as well as patient make-up @frailty @indiantrauma @traumaLMIC @onemilliondead. These differences make a direct transferral of HIC TQIP models onto LMIC irrelevant or misleading and there is a need to develop TQIP assessment applicable to local conditions.

<!-- Ni vil båda flytta fokus från detta, och jag förstår inte riktigt varför det är med kopplingen är luddig. Är det för att säga: det finns TQIPs som ger en fingervisning om hur ICU ska dimensioneras, bestyckas, bemannas och bedrivas Men det finns inga färdiga TQIPs som kan användas för att kvantifiera ICU-bristen i Indien och vi tror det är vanskligt att använda de befintliga eftersom de skapats under så olika förhållanden. Och därför måste vi börja från noll? -->

Knowlege gap
------------
There are compelling reasons to believe there is a shortage of ICU beds in India, but the scope of this shortage as well as the potential for reduction of trauma mortality if this shortage would be remedied is unknown.

WHO and local health care workers believe there is a shortage of ICU beds in Indian hospitals, and that the lives of many patients would have been salvageable if the appropriate ICU resources had been available @ICUbeds. In-hospital trauma mortality in Indian hospitals was double that of high-income countries, and part of this difference could be explained by a shortage of ICU beds in India @indiantrauma.

Aim
----

Quantify the unmet need for ICU beds in tertiary hospitals in urban India.

The theory being examined in this master's thesis is that there exist a subset of the TITCO dataset that did not go to the ICU, but are very similar to the patients that did go to the ICU, in the covariates that caused ICU-treated patients to be admitted. The hypothesis is that these patients were not admitted to the ICU only for lack of ICU resources, not by a medical assessment. 
<!-- Such a group would be akin to an non-experimental control group that PSM can be used to identify and assemble.   -->

If these patients can be identified, that is indicative there exists shortage of ICU beds in tertiary hospitals in India. The size of this group versus the size of the group that did receive ICU care indicate the magnitude of the ICU bed shortage. If, for every patient that did receive ICU care, there was another patient with similar predictors that should hade received ICU care based on our prediction model but did not due to scarcity of ICU resources, that would mean the ICU resources are 50% of what they ought to be.

Methods 
=======

<!-- 
Definitions
------------
An ICU bed is a tiltable hospital bed but included in the term is also staffing and auxiliary equipment. The equipment is to administer oxygen and to monitor vitals. Some ICU beds are equipped with ventilators and if applicable length of ventilation has been registered.

Staffing is one doctor in the ward, and 1:4 nurses per patient but there is not always strict adherence to staffing guidelines. 
If the ICU ward is 100% full, patients might receive ventilation outside the ICU ward. Though the care these patients receive could very well be considered ICU care, they are not registered as such in the TITCO registry.
-->

Study design
------------

<!-- Present key elements of study design -->
Observational retrospective cohort study using data from the Towards Improved Trauma Care Outcomes (TITCO) in India cohort.

Study Setting
-------------

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
Four public university hospitals in urban India are included in the TITCO dataset [@titco]. The participating hospitals were King Edward Memorial hospital (KEM) and Lokmanya Tilak Municipal General Hospital (LTMGH), Mumbai, Jai Prakash Narayan Apex Trauma Center (JPNATC), All India Institute of Medical Sciences, New Delhi and The Institute of Post-Graduate Medical Education and Research and Seth Sukhlal Karnani Memorial Hospital (SSKM), Kolkata. When the data was collected all hospitals were tertiary hospitals. JPNATC had almost 180 beds in a dedicated trauma centre, KEM had no specific trauma ward, LTMGH was a public university hospital with a specific trauma ward with 14 beds and SSKM was a  public university hospital with no specific trauma ward. They are all end-stage public tertiary hospials in their respective areas and there is no possibility to refer patients onward within the public health care system, and ambulance equipment and traffic conditions oftentime makes it medically inadvisable.

The participating hospitals are all public entities, and a majority of patients came from low socioeconomic circumstances in their respective communities.

The data collection was designed in such a way that differences between hospitals would be minimized, but there are some differences in work methods, patient flows and resources that are unavoidable. Triage and patient work flow differed between hospitals. ICU admission criteria were consistent, even if only one hospital had formalized ICU admission criteria. Patients triaged toward the ICU ward were not considered ICU patients, until they had received an ICU bed and had been transferred. Between triage and ICU admission is the resuscitation room. Here - quite simplified - severely injured patients were stabilized before ICU admission and patients with intermediate injuries were kept for observation until most of them were transferred to a ward within one or a few days. The ICU ward occupancy was perpetually at 100 % and so resuscitated patients had to wait for admission with few exceptions. Patients in the resuscitation room were not considered ICU patients even if they were receiving a comparable level of care as ICU patients - patients receiving ionotropic support or patients who were intubated or ventilated. In LTMGH there was no resuscitation room and such a distinction, all admitted trauma patients were immediately considered ICU patients. Therefore for the purpose of this analysis LTMGH patients were considered ICU patients only if they had received ICU care >24 hours.

A dedicated project officer involved in the data collection was on site five days per week, approximately eight hours a day. The project officers covered night, evening and morning shifts in a way so all shift variants were covered every month of collection. Project officers worked in the area where trauma patients arrived to the hospital and collected data for patients arriving during their shift through direct observation. If the officer observed an examination being performed but not documented in the patient records were allowed to ask health care staff for the result. The project officers were strict observers and instructed not to intervene and delay care. If a trauma patient arrived when a project officer was not on site, they were identified from the log books of the nurses on staff in the emergency department and their data was collected from patient records.

The data was collected between July 2013 and December 2015 but the collection period for any individual hospitals did not cover necessarily this whole period.

Participants
------------

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
The eligibility criteria for TITCO were all patients who presented with injuries from trauma at the emergency departments of the participating hospitals with a mechanism of road traffic, railway, fall, burn or assault. Patients who were dead on arrival were excluded, as well as patients who died within

Patients who only had isolated limb injuries were excluded

<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->

<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->

Ethical considerations
----------------------

This study does not require ethical permission, it refers to data collected and shared openly.

Ethical approval for the collection of the data used was obtained from the boards of ethical approval in all four hospitals.

SSKM, LTMGH, KEM and JPNATC name of ethical review boards and approval reference numbers were IPGME&R Research Oversight Committee (IEC/279), Ethics Committee of the Staff and Research Society (IEC/11/13), Institutional Ethics Committee (IEC(I)/OUT/222/14), Institute Ethics Committee All India Institute of Medical Sciences (EC/NP-279/2013 RP-Ol/2013), respectively.

The ethical review boards of all hospitals granted a waiver of informed consent. The trauma patients of in this study included patients in poor physiological, psychological and physical state, oftentimes unconscious or with reduced level of consciousness. The data collected for this study were merely routine data that would have been collected under normal circumstances and did not involve any additional examinations. The collection did not in any way interfere, delay or alter the care provided. Therefore the research team and the ethical review boards concluded that obtaining informed consent would only unneccesarily burden patients and/or family members. Personal details such as names, insurance details or contact information were not recorded.

<!-- Anonymization -->
<!-- ------------- GÄLLER INTE DENNA STUDIE SLAKTA OM BEHÖVS-->
<!-- Hospital identification number -->

<!-- Hospital name has been replaced with a identification number, and the conversion key has not been saved so there is not possible to reverse the conversion. Despite this, a high degree of anonymization can not be expected because the hospitals do have some distinguishable features and these are reflected in the data collected. For example,  70 % of the railroad accidents in the data set is in one of the hospitals, and one of the hosapitals is in close proximity to a big railroad hub.  -->

<!-- Patient identification numbers -->

<!-- Patient names were never recorded, patients were identified by a identification number in the original unencrypted data set. In the anonymization process the identification number has been replaced with another random identification number, and the conversion key has not been saved so there is not possible to reverse the conversion.  -->

<!-- Date observations -->

<!-- Dates recorded in the TITCO-I data set were shifted using a random offset of somewhere between 0 and 5 weeks for each patient, but without losing any relevant information. The offset was designed to not dissipate details that could be of potential use for researchers such as day of week and rough indication of season. The internal relations between dates have been preserved so that the timing of all recordings and discharge in chronological relation to admission are all intact.  -->

<!-- It is important to note that all patients have sequence numbers that corresponded to the chronological order in which they arrived at any of the participating hospitals, and that sequence number has been kept intact. Keeping this sequence number variable intact allows for many useful operations such as examinations of trends in the data, while it is not possible to undo the date randomization by the sequence number variable, such an equation system is still underdetermined. Unfortunately this means it is also not possible to reconstruct the patient volumes at any given hospital on any given day and from that rate of resource engagement. Nor is it possible to distinguish if patients who were made to wait for their first assessment were so because their injuries were deemed as less severe or if it was because health care staff was overwhelmed which is an important distinction. -->

<!-- Age -->

<!-- Very old patients are few, they have been grouped together to prevent identification via the age parameter. -->

<!-- In addition, 47 patient data were randomly deleted to prevent identification via the sequence number variable. -->

<!-- Text-based input -->
<!-- 1% of the text-based inputs in ICD code fields and examination or operation findings were randomly deleted to obstruct identification by this variable.  -->

Variables and data sources/measurements
---------------------------------------

<!--  Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
There are 193 possible observations recorded for each patient. A general presentation of these observations is in table_nums('tab_1')

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("table1_1.rdata")

knitr::kable(table1_1, caption = "Data types collected in the TITCO-I dataset")
```

The quantitative data include recorded parameters such as vital parameters, lab results, Glasgow coma scale (GCS) total score and cumulative time in surgery or ICU, if applicable. Measurement or assessment techniques are standardized.

The qualitative data include categorical data such as coarse categorization or the accident mechanism and injury, details about arrival circumstances, and whether or not a patient had received a medical examination or intervention.

Text data are freehand examination results and operation categorizations, as well as assigned ICD codes. These data are vast and contain a lot of useful information, but also introduce a lot of challenges. These data were not considered for lack of standardization, user dependence, heterogenous grouping caused by using too general ICD codes, and datamining challenges for this type of data.

Bias 
----

<!-- Describe any efforts to address potential sources of bias -->
The TITCO dataset is observational data collected from a clinical environment and not an experimental setting.  Medical interventions, examinations and ICU admission are not randomized, they are a consequence of a patient's medical condition. There exists a causal effect between the independent and the dependent variables: a patient is sent to the ICU precisely because of the severity of their condition. There is also a selection bias in which parameters are recorded and which examinations are performed, introducing detection and observer bias.

There exist a bias in the routines, resources and patient makeup of the participating hospitals as described in chapter Settings. Because they are all tertiary hospitals, about two-thirds of the patients have been referred from other health care providers. This means that there is a selection bias in the patient populations. Further, hospitals with dedicated trauma wards have a propensity to receive more transferred or police-escorted patients.

As stated in chapter Study Settings, most patients have low socioeconomic status. This circumstance doesn't affect the analysis and internal validity, but reduces the generalizability of the results. Most notably, if no ICU bed is available, and it is the attending physicians assessment that it is medically necessary, a patient or their relatives may choose to go to a private hospital. This is economically unfeasible for most of the patients in this study. This bias could not be corrected for, because such patients are pooled in the heterogenous group Discharged Against Medical Advice from where it is not possible to positively identify which patients in this category went on to private care.

Biases like these are expected in data collected in a clinical setting, the analysis model i chosen to adress this. Because the purpose of this thesis is to statistically assess how many patients that should have gone to the ICU but did not, these particular selection biases are not expected to skew the outcome, and therefore do not have to be addressed.

Hospitals with trauma wards could have different admission criteria to the ICU than hospitals with no trauma ward. The same goes for all equipment any hospital has as well as the proficiency of their physicians. Because examining admission criteria to the ICU is part of the analysis, any bias from differences in hospital admission criteria must be minimized. This is why the general linearized model is trained using a training subset and validated using a test subset, as described in the chapter Statistical methods. If this proves insufficient and unsatisfactory it may be necessary to examine and address this bias by comparing ICU admitted patients between hospitals using the same PSM methodology, but this will go beyond the time frame and scope of this thesis.

<!--  
Propensity score matching in itself introduces a source of bias from unfit assumptions. The terminology and potential sources of bias are explained chapter Study Design. In choosing predictor variables the researcher could introduce an overdetermined set and a self-fulfilling prophecy. For this reason the dataset shall be randomly split into two parts of 40% and 60% of the patients. The 60% shall be used to produce the propensity score matching, and the remaining 40% shall be used to verify it. 
-->

<!-- Fråga: vad är distinktionen mellan bias och error med PSM egentligen? är det error eller bias som uppstår om jag ex skulle välja ICU-sjukhuset som prediktor? Är det bias att ha för generösa calipers? Som jag förstår finns risk att få vilket resultat som helst med val metod för matching och limitations ? Och att jag väljer calipers som producerawr resultat som jag tycker känns "rimliga"? Ex jag kanske väntar mig att ICU bristen är 20% för jag föreställer mig att de patienter jag söker är de som är "på ribban" alltså de matchar de "friskaste" 20% av dem som var på ICU. Det känns mindre sannolikt att en patient inte fått ICU om den är jämförbar med någon som var en av de mest skadade 20% på ICU.   -->

Statistical methods
-------------------
<!--

(a) Describe all statistical methods, including those used to control
for confounding
-->
<!-- Propensity score matching (PSM) is useful method for identifying and assembling a control group from within a data set where there is causality between the independent and dependent variables, and a randomized case-control (RCT) study to offset or balance this bias is impractical, immoral or otherwise improper.  -->

The propensity score model is developed using the Training Data and tested using the Test Data. The workflow contains a series of steps. The output of each step is the input for later steps. Steps three and four that involve development and validation of a multivariable regression model have several standard methods to test and improve model validity. By reverting back and adjusting assumptions in an iterative process, the model is continuously improved and the accuracy and validity is increased.

The first step of this analysis is to randomly split the data set into two parts, 60% and 40%. The 60% part is Training Data used to develop the propensity score model, and the 40% part is the Test Part. This is done to detect overfitting of the model to the Training Data.
Training Data is used in steps 2-3 and the resulting model is then validated using Test Data.

Second step is to identify predictor variables: covariates most predictive of the dependent variable, as described in chapter Predictor selection. In this context it means to first identify the conditions and medical interventions received that correlate best to being admitted to ICU care. This selection is based on clinical experience as well as analysis of the Training Data. In the data analysis all properties of the patients that were sent to the ICU will be compared to the patents who did not, in order to identify properties that distinguishes this cohort. Familiarity with the collection of source data is crucial to minimize the risk to choose irrelevant predictors and introduce error.

Third step is to build a multivariable logistic regression model with ICU care as the outcome and the predictors selected in the second step as independent variables. The prediction model is based on the Training Data and predictors are weighed so that the the model awards patients that did go to the ICU corresponding high propensity. This model can then be used to predict new patients propensity to go to the ICU based on their baseline predictor values. By saving 40% in the Validation data set we have a set of patients that is new to the model. This allows to test how well the model correctly predicted ICU patients by awarding them high propensity scores.

Fourth step is to apply this model on patients from the Test Data. Using the model, the standard variable propensity score is calculated for each patient in the Test Data set. The propensity score determines the likelihood that a patient will receive ICU care, given their condition. If the model is suitable, it will correctly identify patients that did go to the ICU by awarding them high propensity scores, but non-ICU patients that were similar to this cohort will also get high propensity scores.

<!-- 
fråga: vi är ju ute efter ge vissa non-ICU pat höga propensity scores för att sedan kunna identifiera och filtrera fram dem i propensity score matchingen steg 5. Men hur identifierar man skillnaden mellan en patient som förmodligen borde fått ICU, och en patient modellen felaktigt spottat ut?
 -->
One standard method to evaluate how well a classification model performs in making binary predictions is apply it on new data and perform an Area Under Curve (AUC) calculation. In this calculation, the rate of true and false positives and negatives are calculated for each cut-off, which produces a curve as below. If the model has a 100% prediction accuracy, both sensitivity, specificity and AUC are all maximized at 1, the red lines. If the model has just as much ability to predict the outcome as a cointoss then sensitivity, specificity and AUC are 0.5, the blue line. For the purposes of this thesis, neither of these outcomes are desirable, and any model will perform somewhere between these extremes.

If there is no preference between rate of false positives and negatives, Youden's index is the optimal cut-off in propensity score with the best sensitivity and specificity. Patients that have a propensity score higher than this optimal cut-off in propensity score are patients that according to the model should have received ICU treatment. This number of patients can then be put in relation to the patients who received ICU treatment to calculate the unmet need of ICU beds.

<!-- peta in plot roc_test -->
![Exempel ROC kurva](C:/Users/maria/Desktop/stash/test_ROC.png)
Fifth step is bootstrapping step three and four to produce a confidence interval for the predictions of unmet ICU meds. Bootstrapping is a method of randomly resampling from the same data set and produce an output with some spread, a mean, standard deviation and 95% confidence interval. This will produce several samplings from which to produce several general linearized models, several AOC curves, several Youden's cut-offs, several estimates of patients whose propensity scores are higher than this cut-off and from this several proportions of unmet ICU bed needs. Several estimates with a mean, standard deviation and 95% confidence interval within which the best estimate of the unmet need is.

Bootstrapping is preferred over propensity score matching because of the weaknesses of the propensity score matching method and the available data size. 
<!-- återkom jag gissar bara -->

<!-- Fifth step is PSM. In PSM members of the subset that did not get ICU care are matched to patients who did go to the ICU, matching is based on patient propensity scores. There are a few different ways to perform this matching but they are all imperfect because propensity scores are unique and perfect matching is unattainable. There are control mechanisms and limitations but a risk of applying too fix or too loose matching criteria and inadvertently vastly under- or overestimate the  need of ICU beds remain.   -->

<!-- (b) Describe any methods used to examine subgroups and interactions -->

The two subgroups were patients admitted to the ICU, and the patients who were not. Admission to the iCU is a binary yes or no, length of stay has not been taken into Consideration.

Missing Data
-------------
The data completion rate is presented in table_nums('tab_1').

Rstudio default behaviour when making a multivariable regression model is to drop all entries with NA for predictors. This reduces the sample size, power and ultimately quality of the analysis. For the purposes of this statistical analysis, missing data in qualitative varibles is consdidered its own category. For quantitative data any variable with more than 10%  missing values are converted to an integer categorical variable and missing values treated as its own category. For systolic blood pressure and respiration rate, the interval brackets of the Revised Trauma Score was used, and for heart rate and blood oxygen saturation the brackets of National Early Warning Score (NEWS) was used. NEWS take into account whether or not the patient was receiving oxygen at the time of the recording.

Zero values for variables where it is possible to have zero values were left as is, such as respiration rate and heart rate if pulse was also zero and vice versa. However, when zero values are implausible, like when either heart rate or pulse were zero but the other was not, the observation is discarded. Data that were questionable such as when GCS observations did not add up were discarded.

<!-- FRÅGA: jag vill droppa noll i SC, eftersom noll är orimligt. men då droppar jag ca 3000 observationer på en variabel som inte ens har en stjärna i sginifikans för GLM. Droppa sc som prediktor alltogether? yes droppa sc
-->

<!-- (d)
Cohort study: If applicable, explain how loss to follow-up was addressed

Case-control study:If applicable, explain how matching of cases and
controls was addressed

Cross-sectional study: If applicable, describe analytical methods
taking account of sampling strategy

(e) Describe any sensitivity analyses
-->

Predictor selection 
-------------------

Predictors are independent variables with strong correlation with the dependent variable. The dependent variable is ICU admission so the predictors are variables indicative of how badly injured the patient is. For the purpose of this study, causality and chronology is unimportant, so variables such as patient survival which is a strong indicator of injury severity, could be included as a predictor.

Predictor variables will be selected using both clinical judgement and data analysis, where clinical judgement is superior. Variables deemed relevant from clinical experience don't need to present with convincing statistical relevance from the data analysis, but variables with convincing stastistical relevance from the data analysis still need to make sense from clinical judgement to be included. 
<!-- Indian clinicians experience may need to be applied because the data analysis might output a variable as relevant but the reason may not be obvious to someone working outside the Indian health care section. Many relevant variables for predictors can be intercepted in this way. -->
The TITCO-I data set is analysed mathematically in pursuit of suitable predictor variables. The data is divided into two subgroups, the cohort that were admitted to the ICU and the cohort that was not. By comparing these subgroups, variables that distinguishes the former from the latter are identified.  
Mean and standard deviation for the quantitative data is analysed for the ICU and non-ICU cohort respectively. The cohorts are compared and p- and t-values are calculated for each variable.

Based on clinical experience, new variables can be constructed out of the original data, for consideration of their predictor potential. Second and third delays, corresponding to the delays between incident and hospital admission and between admission and first survey have been proposed to have high relevance for patient outcome @indiantrauma.

The free hand text-based information available contain information about examination findings, surgery types and ICD codes. Many of these data fields contain crucial information about patient condition, but also present a few challenges as described in chapter: Variables and data sources/measurements, which is why they fall outside the scope of this thesis.

<!-- It is also plausible to include aggregate variables, by combining unrelated variables of the same type or across varaible types. Selection and construction of such aggregate parameters to be explored will be based on clinical experience not data analysis. An example of this is the heterogenous qualitative parameter "taken to the operation theatre". Many operation types are strongly correlated to ICU admission - especially if they were performed within the first hour of admission - but most of the operations performed are lower priority orthopaedic operations with no correlation to ICU admission. There is a need to combine operation timing and free text fields "operation type" and "intraoperative findings" to filter out operations with higher propensity for ICU admission. -->

Epi::ci.lin(glm_model)

Results
=======

Missing data
------------
Återkom: XX patients were discarded because of missing data.

Predictor selection
-------------------
<!--
ingenstans egentligen ska du beskriva metoden, men isåfall ska beskrivning och motivering av metodval finnas i metoddelen
inga beskrivningar i resultatdelen.
-->

VARIABLES BASED ON CLINICAL EXPERIENCE
Clinical experience compels inclusion of vital parameters. Such variables are GCS, heart rate, systolic blood pressure and respiration rate that were documented twice for each pateint, first and second survey. These variables differ statisticvally significant between the ICU and non-ICU cohort, see table_nums('tab_2') and table_nums('tab_3').  
Because ICU admission for GCS was better correlated to low scores in either test eye, motor or verbal than total GCS, GCS scores is translated to eleven categorical variables for each survey with GCS "1", "1s", "2" and "other" for all three tests, where "1s" represent when GCS test couldn't be performed. For the verbal test 1v is given when the patient intubated, and for the eye test 1c is given when the eyes are swollen shut.

Surgical airway, intubated  and lov greater than 0, intercostal drain within an hour
<!-- FRÅGA MÅNDAG: actutally ICU admission not 100% for any of these. Did a trial with and without how should I verify if it was good
saw_1 rimligt pga ca 76% ICU. aggregerat before_arrival och within 1 hour? 
jag har också noterat att vid 4-5 reps som är en rimlig mängd iterationer svingade det som fanken med lov, ej med saw. fått in intub + icd_1, incl (doctors assessment on arrival). yes alla utom lov. inkludera motivering i koden.
-->

<!-- cleaning, vad är orimligt? tubad utan lov? tubad utan licu? -->

<!-- hur får man in GLM2 objekt? börja googlingen med vad? also glm summary ganska påtaglig skillnad mellan glm för traindatan vs glm för hela datan SVAR: du har inte missing i din predictors modell ju. -->

In addition to these, significant differences in ICU admission were expected from the categorical variables "mode of injury" and "mode of transportation", and whether or not the patient had been transferred from another health care provider.

Mode of injury was included based on clinical experience because different modes of injury have significantly different impact forces and therefore significant differences in how injured the average patient is. Traffic accidents make up ~40% of the patients in the TITCO-1 study, and this category is further divided into "bicyclist", "driver", "passenger", "motorcyclist",  "pedestrian" and "unspecified".
One notable example are railroad accidents that are high-impact severe accidents with high mortality, and therefore high likelihood of ICU admission. One common type of accident in this category is falling out of a moving train car, and although data wasn't collected for patient income, this is a type of accident which affect the  the low-cost general train cars and almost exclusively affect low-income patients.
Regarding mode of transportation it is important to present some local circumstances that determined the categorization. In India pre-hospital triage and care is fairly uncommon for trauma patients, and many patients arrive to hospitals by taxis or private cars because it is faster @goodsamaritan. Ambulances are mostly used for interhospital transportation, but ambulances are stationed at places where injuries are common, such as major railroad road traffic hubs. 
The TITCO-1 data was collected 2013-2015, before the enactment of Sec 134A- (1), colloquially known as "The good samaritan law" in 2016. This law protects any citizen that brings a patient to the hospital from civil and criminal liability, and was enacted to increase the prportion of patients that reveive care within the "golden hour" and reduce pre-hospital mortality. By-standers were often reluctant to help accident victims, fearing legal intervention. For this reason, many badly injured patients were brough in by police @goodsamaritan. Therefore, when mode of transportation is included as a predictor, the seven categories of the TITCO-1 data "Ambulance","Carried by man","Other","Police" "Private car","Taxi, motor rickshaw" were reduced to three: "Police", "Ambulance directly from the injury" and "Other".

Whether a patient had been transferred from another health care provider or not is also selected as a predictor. A transferred patient with unstable vitals at the same level as a directly admitted patient is likely worse injured, because the first hospital has been unable to stabilize the patient. Being as the participating hospitals are all tertiary hospitals, 70% of the patients had been transferred, and the proportion varies btween 50 and 90% between the centres.

<!-- fråga: är det här pa a) fel ställe? Det är en motivering till de kliniska avvägningarna och hör hemma i "resultat" men de ör ocskå ganska bakgrundsmässiga b) baserat på Anna, Manor och Kapil och så har jag hittat en studie som talar om samma fenomen -->

As demographic variables, sex and age were included.

VARIABLES BASED ON DATA ANALYSIS

The quantitative data results see table_nums('tab_2')

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("table5.rdata")

knitr::kable(table5, caption = "Means, standard devations of quantitative variables of ICU and non-ICU admitted patients, respectively and p- and t-values of comparison between the sets")
```

For the qualitative data, the propensity of each parameter for ICU admission is compared to the proportion of patients that were admitted to the ICU from the whole TITCO-I data set, about a third. If considerably more than a third of each qualitative parameter are admitted to the ICU, then that parameter is of interest. It has been relevant to combine qualitative parameters into aggregate qualitative variables. For example, the variable intubation at 1 hour has two options indicating the patient is intubated: "before arrival" meaning the patient was intubated at a their first health care facility and "yes" meaning the patient was intubated within the first hour. Because these two options both mean the patient was intubated after 1 hour, an aggregate parameter has been created for this. Because the data set also contain information of whether a patient was intubated within the first and 24th hour the parameter "intubated after 24 hours" were created that summarize the patients who were intubated on arrival, the patients who were intubated within the first hour and the patients who were intubated between 1 and 24 hours.

The proportions of the qualitative data are presented in

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("qualitativetable.rdata")

knitr::kable(qualitativetable, caption = "Proportion of patients admitted to the ICU for all variants of all quuantitative variables")
```

From table_nums('tab_2'), the vital parameters selected from clinical experience also present with a statistically significant difference between the ICU and non-ICU patient cohorts.

The data analysis show significantly higher ICU admission rates for patients with high New Injury Severity Score, patients that had received ≥ units of blood, patients that were sent to the operation theatre within the first hour of admission, patients with serum creatinine over 3 mg/dL, and patients that dies, and these variables are included as predictors.

<!--

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->

You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this: `r 2+5`.

Discussion
==========

<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias
-->

In order to limit the scope of this thesis work, a few quite relevant factors could not be considered. One of the most important factors was temporality. Emergency rooms everywhere face significant variations in patient loads over the week or special occasions such as the Diwali festival. A different approach completely but to the same effect would have been to consider the ICU wards at the participating hospitals and chronologically determine which patients did not receive ICU care because these wards were at capacity. Though intuitively it would seem possible to do given the TITCO-I dataset, it would have been very difficult to do practically.

Length of stay in hours is ICU care was recorded in the TITCO-I data set, but for the purpose of this thesis ICU admission has been reducedn to a binary 'yes' or 'no'. Because it is not uncommon to discontinue ICU care prematurely in favor of patients with more dire needs, the next step would be to consider such circumstances.

<!--
6 hours med eller ej? 
-->

<!--
Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->

Conclusion
==========

<!-- Give a short conclusion, corresponding to the aim -->

References
==========

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->